---
title: "Hello, World!"
date: "2022-08-01"
categories: [Miscellaneous]
toc: true
code-fold: false
---

This first post is to check whether the features I want for this blog work as desired.

These can be summarized in a non-exhaustive list as: 

- math
- R / Python code
- interactive graphs (plotly, leaflet, ...)
- syntax highlighting
- markdown
- ...

## Math

The OLS estimator is given by the equation $\hat\beta_\text{OLS} = (X^\mathsf{T} X)^{-1} X^\mathsf{T} y$.

On the other hand, the _ridge_ estimator is given by the following formula

$$\hat\beta_\text{ridge} = (X^\mathsf{T} X + \lambda I)^{-1} X^\mathsf{T} y$$

where $\lambda \in [0, \infty)$ controls the amount of shrinkage applied to the coefficients.

## R code

```{r}
library(dplyr)
library(ggplot2)

p <- iris %>% 
  ggplot(aes(x = Petal.Length, y = Petal.Width, color = Species)) + 
  geom_point() + 
  theme_classic()
```

vanilla ggplot

```{r}
plot(p)
```

plotly plot

```{r}
plotly::ggplotly(p)
```
